Follow-up Samples

### 2017-01-11 submission/manuscript_outline_20161024.Rmd
update background of abstract

**Background:** Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease. Previous research suggests a link between stool bacterial microbiome and CRC.   In this study, we used a model based on the microbiome, demographics, and prior medical history to classify individuals as having a lesion (i.e. adenoma or carcinoma).  We then used this model to characterize the change in the gut microbiota before and after surgery.  The overall objective was to investigate the changes in the microbiome after surgery in patients with adenomas or carcinomas.  This model was tested on a `r sum(initial_cm$ref_yes)` person group that included samples before and after treatment to allow for the assessment of how the model adjusts risk after treatment. 
**Results:** The model used for prediction had an AUC of `r format(model_test_AUC_summary["full", "AUC"], digits = 3)`. For the follow up samples our Random Forest model significantly decreased the positive probability of lesion compared to the initial samples for both adenoma (P-value = `r format(IF_model_wilcox_paired_summary['all_adenoma', 'BH_correction'], scientific = TRUE, digits = 3)`) and carcinoma (P-value = `r format(IF_model_wilcox_paired_summary['carcinoma_only', 'BH_correction'], scientific = TRUE, digits = 3)`).  Our model predicted that `r format((followup_cm["pred_no", "ref_no"]/(sum(followup_cm$ref_no) + sum(followup_cm$ref_yes)))*100, digits = 3)`% of the 67-person cohort had normal colons and `r format(((followup_cm["pred_yes", "ref_no"] + sum(followup_cm$ref_yes))/(sum(followup_cm$ref_no) + sum(followup_cm$ref_yes)))*100, digits = 3)`% had a lesion.  Some OTUs that changed the most before and after treatment included OTUs that were affiliated with members of `r pvalue_common_imp_vars[order(pvalue_common_imp_vars$Pvalue), "Genus"][1:2]` and `r pvalue_common_imp_vars[order(pvalue_common_imp_vars$Pvalue), "Genus"][3]`. 
**Conclusions:** Our model suggests that treatment does significantly reduce the probability of having a colonic lesion.  Further surveillance of these individuals will enable us to determine whether models such as the one we present here can also be used to predict recurrence of colorectal cancer.

>>>>

**Background:** Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease. Previous research suggests a link between stool bacterial microbiome and CRC.   In this study, we used a model based on the microbiome, demographics, and prior medical history to classify individuals as having a lesion (i.e. adenoma or carcinoma).  We then used this model to characterize the change in the gut microbiota before and after surgery.  The overall objective was to investigate the changes in the bacterial microbiome after surgery in patients with lesion.  Specifically, we wanted to identify what within the community was different within those undergoing surgical removal of lesion. 
**Results:** The model used for prediction had an AUC of `r format(model_test_AUC_summary["full", "AUC"], digits = 3)`. For the follow up samples our Random Forest model significantly decreased the positive probability of lesion compared to the initial samples for both adenoma (P-value = `r format(IF_model_wilcox_paired_summary['all_adenoma', 'BH_correction'], scientific = TRUE, digits = 3)`) and carcinoma (P-value = `r format(IF_model_wilcox_paired_summary['carcinoma_only', 'BH_correction'], scientific = TRUE, digits = 3)`).  Our model predicted that `r format((followup_cm["pred_no", "ref_no"]/(sum(followup_cm$ref_no) + sum(followup_cm$ref_yes)))*100, digits = 3)`% of the 67-person cohort had normal colons and `r format(((followup_cm["pred_yes", "ref_no"] + sum(followup_cm$ref_yes))/(sum(followup_cm$ref_no) + sum(followup_cm$ref_yes)))*100, digits = 3)`% had a lesion.  Some OTUs that changed the most before and after treatment included OTUs that were affiliated with members of `r pvalue_common_imp_vars[order(pvalue_common_imp_vars$Pvalue), "Genus"][1:2]` and `r pvalue_common_imp_vars[order(pvalue_common_imp_vars$Pvalue), "Genus"][3]`. 
**Conclusions:** Our model suggests that treatment does significantly reduce the probability of having a colonic lesion.  Further surveillance of these individuals will enable us to determine whether models such as the one we present here can also be used to predict recurrence of colorectal cancer.
